---
title: "Assesment 2"
author: "Malgorzata Olesiewicz"
date: "31 January 2019"
output: pdf_document
---
---
title: "Microeconometrics 2018/2019 - 1st Assignment"
date: "20th of December, 2018"
output: 
  pdf_document:
    latex_engine: xelatex
fontsize: 11pt
---

#### Instructor: Tomas Repasky
#### Group Members: 
- Johannes Wagner  
Humboldt-Universitat Berlin - ID 598797 - Msc Statistics  
<wagnejoh@hu-berlin.de>
- Malgorzata Paulina Olesiewicz  
Humboldt-Universitat Berlin - ID 598939 - Msc Statistics  
<malgorzata.paulina.olesiewicz@student.hu-berlin.de>

- Kevin Hoppe  
Humboldt-Universitat Berlin - ID 598247 -  Msc Statistics  
<Kevin.Hoppe1@web.de>  

## Approximate individual contributions: 
```{r, echo = FALSE}
suppressWarnings(library("kableExtra"))

a=c("1",
    "2",
    "3",
    "4") 
   

b=c("70% progr. + 50% theory",
    "70% progr. + 70% theory",
    "70% progr. + 30% theory",
    "33% theory") 

c= c("-", 
     "-",
     "40% theory",
     "33% theory") 
 
d= c("30% progr. + 50% theory", 
     "30% progr. + 30% theory",
     "30% progr. + 30% theory",
      "30% progr. + 30% theory")

work = data.frame(a,b,c,d)
colnames(work)= c("Task","Johannes","Malgorzata","Kevin")
kable(work)%>% kable_styling() 
```


$$
\pagebreak
$$


```{r setup, include=FALSE}
### Setting up the enviroment
library(foreign, quietly = TRUE)
library(foreign, quietly = TRUE)
library(margins, quietly = TRUE)
library(mfx, quietly = TRUE)
library(caret,quietly = TRUE)
```

```{r}
### loading the data and looking up meaning of the variables

data=read_dta("loanapp1.dta")
(variables=cbind(colnames(data),attr(data, "var.labels")))
names(data)
head(data)
```

##Chosing the variables for the model

To create the logit model we have chosen fallowig variables: 

1. Gender (gender)
2. Loan amount (loanamt)
3. Married (married)
4. Monthly income (appinc)
5. Morgage history (mortno)
6. Net worth (netw)
7. Age (old) 



```{r}
model=glm(data$approve~data$gender+data$loanamt+data$married+data$appinc+data$mortno+data$netw+data$old, data=data, family = "binomial")
summary(model)
```

###Results interpretation


###Sensitivity and Specificity

