---
title: "Microeconometrics Assignment 2 - Task 4"
author: "KH"
date: "3 Februar 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

setwd("C:/Users/Nutzer/Desktop")

## Packages:

require(foreign)
require(margins)
require(mfx)
require(dplyr)
require(tidyverse)
require(nnet)
require(mlogit)

```



```{r, include=FALSE}

## Load data and extract the 7 variables that we chose 
## as a group:
##
## 1. Education (sch): [0,1]
## 2. marital status (married): [0,1]
## 3. Existence of private mortgage insurance (inson): [0,1]
## 4. Credit History (cons): [1,2,3,4,5,6]
## 5. Former bankruptcy (pubrec): [0,1]
## 6. Ethnic majority (white): [0,1]
## 7. Ratio of income and liabilities (obrat) : [0-95]

dat.raw <- read.dta("loanapp1.dta")
dat.red <- select(dat.raw, approve, sch, married, inson, cons, pubrec, white, obrat); as.tibble(dat.red)

## Coding categorial variables as dummy variables:

dummy.var <- function(var){
  
  if(min(unique(var), na.rm = T) == 0){
    var <- var + 1
  }
  
  len.var <- length(var)
  levels.var <- var %>% as.factor %>% levels %>% length
  levels.dummy <- levels.var - 1
  
  dummy <- matrix(NA,
                  nrow = len.var,
                  ncol = levels.dummy)

  for(j in 1:levels.var){
    
    match.loc <- which(var == unique(var)[j])
    
    for(i in match.loc){
      
      dummy.vec <- rep(0, levels.dummy)
      dummy.vec[unique(var)[j] - 1] <- 1
      dummy[i,] <- dummy.vec
    }
  }
  return(dummy)
}

sch <- dummy.var(dat.red$sch); head(sch)
married <- dummy.var(dat.red$married); head(married)
inson <- dummy.var(dat.red$inson); head(inson)
cons <- dummy.var(dat.red$cons); head(cons)
pubrec <- dummy.var(dat.red$pubrec); head(pubrec)
white <- dummy.var(dat.red$white); head(white)

## Put together data set for regression models:

dat <- data.frame("appr" = dat.red$approve,
                  "educ" = sch,
                  "marry" = married,
                  "insur" = inson,
                  "credithist" = cons,
                  "bankr" = pubrec,
                  "white" = white,
                  "oblig" = dat.red$obrat)

dat.mnl <- mlogit.data(dat, shape = "wide", choice = "appr")

## Formula for regression model:

form <- 'appr ~ educ + marry + insur + credithist.1 + credithist.2 + 
         credithist.3 + credithist.4 + credithist.5 + bankr + white + oblig'

form.mnl <- mFormula(appr ~ educ + marry + insur + credithist.1 + credithist.2 + 
                       credithist.3 + credithist.4 + credithist.5 + bankr + white + oblig)

## Estimate a binomial logit model:

head(dat); binom <- glm(formula = form, data = dat, family = "binomial"(link = "logit"))
summary(binom)

## Estimate a multinomial logit model:

head(dat.mnl); multinom <- multinom(formula = form, data = dat)
summary(multinom)

## Comparison of coefficient estimates:

beta.diff <- cbind(coef(binom) - coef(multinom)); beta.diff

```


The difference in the coefficient estimates is zero when rounded to even the fourth 
decimal place:

```{r, echo = FALSE, include = TRUE}

beta.diff


```


This is because the multinomial logit model reduces to the 
binomial logit model in case of a binomial dependent variable as can be seen in the formulas. In the multinomial logit model, the probability $\pi_{ij}$ of individual $i$ choosing alternative $j$ is given by:

  $$ \pi_{ij_{multinomial}} = \frac{\exp(x_{i}'\beta_{j})}{\sum_{r=1}^J \exp{x_{i}'\beta_{j}}}. $$

Compare this to the binomial logit model, where the probability $\pi_{i}$ of individual $i$ 
picking alternative $j = 1$ is given by:

  $$ \pi_{i_{binomial}} = \frac{\exp(x_{i}'\beta)}{1 + \exp(x_{i}'\beta)}.  $$
  
In the multinomial model, due to identification constraints, $\beta_{1}$ is fixed at $0$. This establishes $j = 1$ as the reference category. For the remaining $J-1$ categories, $\beta_{j}$ coefficients are estimated. If the dependent variable has only two categories, i.e. $J = 2$, this means that only one $\beta$ and one $\pi_{i}$ need to be calculated (for the one category that is not the reference category) so the index $j$ in $\pi_{ij}$ and $\beta_{j}$ can be dropped. Since the constraint for $\beta_{1} = 0$ means that $\exp{(x_{i}'\beta_{1})}$ evaluates to $1$, this reduces the denominator in the multinomial logit model to $1 + \exp(x_{i}'\beta_{2})$. After dropping the now obsolete index of the coefficient vector, the two formulas given above are equal.


